Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages

### 2021-05-28 paper/manuscript.Rmd
Write introduction

---
title: "More than space for time: revealing the tight links constraining the temporal dynamics of network structure"
author: "Alain Danet, Maud Mouchet, Loubna El Madouri, Colin Fontaine (ordre non fixé)"
self_contained: yes
date: "`r Sys.Date()`"
tables: yes
bibliography: references.bib 
fontsize: 12pt
output:
  bookdown::html_document2:
    fig_caption: true
    theme: "journal"
---

@@

theme_set(mytheme)
bookdown::html_document2
library(bookdown)
```
```{r}
net_var_bm <- c("log_rich_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
net_var_rich <- c("log_bm_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
```

@@

```{r}
loadd(com_analysis_data, full_data2)
```
```{r}
nb_st <- unique(full_data2$station) %>% length
```

@@

With the goal to characterize the links among the temporal variations of network
and community structure and the temporal variations of community biomass and
species richness, we characterized the temporal trajectories of network and
community structure at each site. We used a method based on second order
polynomial which allows to assess both the direction and the monotony of
temporal trends [@rigal_method_2020]. To avoid complex variations of biomass and
species richness, we selected the stations that displayed monotonic trends for
theses variable. Over the `r nb_st` stations, the selection procedure resulted
in XX stations for biomass and species richness analysis.

@@

@@

- Main results:
  - different mechanisms for richness and biomass
  - space for time subtitution works for biomass and tlvl but not for
    connectance

@@

To the difference with community biomass, the temporal trends evenness of
biomass and individual were positively related to the temporal trends of species
richness, meaning that as community biomass temporally increase, the
distribution of biomass and abondance become more even. Our results indicate
that biomass trends are linked to few big individuals while richness trends are
linked to overall shift in dominance in communities.

Surprinsingly, we found the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical foodweb
[@dunne_network_2006]. However, the relationship become slightly negative for
species poor communities. An explanation might be that species-rich communities
already contains a significant number of piscivores, which have the most links
whereas species poor communites contains few piscivores.

@@

Spatially and temporally, the species richness was positively linked to that of
biomass. Spatially, this result is in accordance with numerous studies was
already found in streams [@woods_testing_2020; @danet_species_nodate] but also
in grasslands, salt marches and forests [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018; but see
@grace_integrative_2016]. Therefore, our results confirm that BEF prediction
works also in time. The positive relationship of species richness on total
biomass might be due to resource use complementarity [@carey_determining_2011],
which is held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability of getting high biomass
species with higher species richness [@loreau_biodiversity_2001].

@@

The temporal trend of the average trophic level had no relationship with the
one of species richness. It is surprising because one can imagine that loss and
gain of species would reveal assembly/disassembly mechanisms in terms of trophic
structure. However, we found that when species richness increases, the
distribution of biomass and individuals among species become more even in the
community, meaning that dominance decrease. Species richness has often been
found to be positively or negatively related to individual eveness
[ @maureaud_biodiversityecosystem_2019].

@@

We found that community biomass was linked positively related to species richness
both in space and in time, indicating that space for time might be use. In the
same way, we found a positive positive relationship between average trophic
level and community biomass both in time and in space. However, it was not the case
for connectance and average trophic level that were respectively weakly or not
linked to species richness in time whereas our result shows a strong negative
relation in space. Moreover, connectance displayed a strong positive
relationship (move to results?) with biomass in time but not in space.

Space for time substitution is a widely
used tool in ecology [@pickett_space-for-time_1989; @blois_space_2013;
@damgaard_critique_2019]. However, the drivers of spatial variation might not
be the ones that drive community variation in time. The successional processes
driving the assembly or the disassembly of communities might not be reflected in
spatial gradient.

@@

We showed that BEF works in time but that richness and biomass temporal trends through
time relates on different ecological processes inside communities. We further
found that space for time subtitution does not always work.

@@

sp_bm <- sp_relation_plot2 %>%
  filter(x == "log_bm_std")
p_sp_bm <- sp_bm$gg 
names(p_sp_bm) <- sp_bm$y

reorder_p_sp_bm <- p_sp_bm
reorder_p_sp_bm[[1]] <- p_sp_bm$ct_ff 
reorder_p_sp_bm[[2]] <- p_sp_bm$w_trph_lvl_avg 
reorder_p_sp_bm[[3]] <- p_sp_bm$piel_nind 
reorder_p_sp_bm[[4]] <- p_sp_bm$piel_bm 
reorder_p_sp_bm[[5]] <- p_sp_bm$log_rich_std 
```

>>>>

---
title: "Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages"
author: "Alain Danet, Maud Mouchet, Loubna El Madouri, Colin Fontaine (ordre non fixé)"
self_contained: yes
date: "`r Sys.Date()`"
tables: yes
bibliography: references.bib 
fontsize: 12pt
output:
  bookdown::html_document2:
    fig_caption: true
    theme: "journal"
---

@@

theme_set(mytheme)
```
```{r}
net_var_bm <- c("log_rich_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
net_var_rich <- c("log_bm_std", "ct_ff", "w_trph_lvl_avg", "piel_bm",
      "piel_nind")
```

# Introduction

## community structure are dynamics and affect ecosystem functions

Biodiversity to affect ecosystem properties such as total biomass, productivity
and stability [@tilman_biodiversity_2014; @loreau_biodiversity_2001]. Numerous
studies have shown the decline of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as ongoing extinction [@barnosky_]. Consequences of such trends on various
aspects of community structure are not well known whereas numerous ecological studies
have shown that the different aspects of community structure are tightly linked [REF].

The link between Biodiversity and ecosystem functioning is one of the most
famous example, experimental studies are in accordance to show a positive
relationship between species richness and community biomass
[@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But this relation is
not always found in natural settings [@grace_integrative_2016;
@li_relationship_2018; @maureaud_biodiversityecosystem_2019]. Also, numerous
studies have found relation between species richness and evenness, showing that
communities become more or less even as species richness increase
[@soininen_relationship_2012; @ma_species_2005;
@maureaud_biodiversityecosystem_2019], highlighting the diversity of
assembly processes in ecological communities.

The distribution and the number of ecological interactions, i.e. the structure
of ecological networks, also covary. In food-web, connectance, which measures the
degree of generalism in a network, has been theoretically shown to theoretically
positively scale with community biomass [@thebaut_food-web_2003]. Connectance
has been also shown to be negatively related to species richness in empirical
food-webs [@dunne_network_2006]. Trophic level has also been shown to be
positively linked to total biomass in theoretical models
[@wang_biodiversity_2018]. All together, these studies show that the
different dimensions of network structure are closely linked.

The space for time substitution allows to use the studies mentioned above to
predict how changes in species richness or community biomass will in turn affect
community and network structure. The hypothesis that space can be use to
surrogate space is often implicitly made in ecological studies. This method
began to be widely used in the early 20th century to infer the temporal
succession of vegetation from space variations [@pickett_space-time_1989].  It
supposes a homology between variations in space and in time i.e. that spatial
sequence corresponds to temporal sequences. This methods have raised criticisms
[@damgaard_critique_2019], with studies showing the coherent in space and time
[@blois_space_2013] or not [@johnson_testing_2008; @adler_contrasting_2007].
Spatial turnover of species composition, dispersal limitation, disturbance and
stochastic processes  has been hypothesised to be responsible for the failure of
space for time substitution [@adler_contrasting_2007; @blois_space_2013].

In this line, how biomass and species richness temporal trends are linked to
change in food-web structure is still an open question, as well as if space for
time substitution works in food-web assembly. The increasing availability of
ecological temporal series makes possible to assess those questions.

Using a standardized monitoring of fish freshwater communities on the period
1995-2018, we assessed the relationships among temporal trends of food-web
structure, trends of community biomass and trends in species richness.  
Moreover, we tested if the those relationships differs between communities that
lossed or gained biomass/species richness across time, reflecting potential
difference between disassembly and assembly processes. We finally assessed if
spatial relationships display the same patterns than temporal ones, to assess
the validity of space for time substitution.

@@

```{r}
loadd(com_analysis_data, full_data2, st_mono_trends_combined_list)
```
```{r}
nb_st <- unique(full_data2$station) %>% length
nb_st_var <- map_dbl(st_mono_trends_combined_list$station, length) 
names(nb_st_var) <- st_mono_trends_combined_list$variable
```

@@

With the goal to characterize the links among the temporal variations of network
and community structure and the temporal variations of community biomass and
species richness, we characterized the temporal trajectories of network and
community structure at each site. We used a method based on second order
polynomial which allows to assess both the direction and the monotony of
temporal trends [@rigal_method_2020]. To avoid complex variations of biomass and
species richness, we selected the stations that displayed monotonic trends for
theses variable. Over the `r nb_st` stations, the selection procedure resulted
in respectively `r nb_st_var["log_bm_std"]` stations  and `r nb_st_var["log_rich_std"]`
for community biomass and species richness analysis.

@@

In order to control for the uncertainties of the estimated temporal trends, we
weighted each temporal trends by the inverse of the product between the standard
error of the estimated trends, i.e. slopes, for the response and the explicative
variable.

@@

@@

To the difference with community biomass, the temporal trends evenness of
biomass and individual were positively related to the temporal trends of species
richness, meaning that as species richness temporally increase, the
distribution of biomass and abondance become more even. Our results indicate
that biomass trends are linked to few big individuals while richness trends are
linked to overall shift in dominance in communities.

Surprinsingly, we found the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical foodweb
[@dunne_network_2006] and also in the present study. This result suggests that
the structure of food-webs might be not very sensitive to variation of species
richness and that species loss and gain concerns species of low trophic level,
which in turn does not affect connectance. However, the relationship become
slightly negative for species poor communities. An explanation might be that
species-rich communities already contains a significant number of piscivores,
which have the most links whereas species poor communites contains few
piscivores.

@@

Spatially and temporally, the species richness was positively linked to that of
community biomass. Spatially, this result is in accordance with numerous studies was
already found in streams [@woods_testing_2020; @danet_species_nodate] but also
in grasslands, salt marches and forests [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018; but see
@grace_integrative_2016]. Therefore, our results confirm that BEF prediction
works also in time. The positive relationship of species richness on total
biomass might be due to resource use complementarity [@carey_determining_2011],
which is held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability of getting high biomass
species with higher species richness [@loreau_biodiversity_2001].

- Add loss of species richness, loss of biomass and increase dominance?

@@

---
#The temporal trend of the average trophic level had no relationship with the
#one of species richness. It is surprising because one can imagine that loss and
#gain of species would reveal assembly/disassembly mechanisms in terms of trophic
#structure. However, we found that when species richness increases, the
#distribution of biomass and individuals among species become more even in the
#community, meaning that dominance decrease.
---

@@

We found that community biomass was linked positively related to species richness
both in space and in time, indicating that space for time might be use. In the
same way, we found a positive relationship between average trophic
level and community biomass both in time and in space. However, it was not the case
for connectance and average trophic level that were respectively weakly or not
linked to species richness in time whereas our result shows a strong negative
relation in space. Moreover, connectance displayed a strong positive
relationship with biomass in time but not in space. This result is in accordance
with other studies which found mismatch between spatial and temporal patterns
[@adler_contrasting_2007; @johnson_testing_2008]. This mismatch is mainly due to the crucial assumptions
that variations in space will correspond to differences of assembly stages and
then reflect the assembly/disassembly trajectories in time of communities [@].

@@

We showed that BEF works in time but that richness and biomass temporal trends
through time relates on different ecological processes. We further found that
space for time subtitution does not always work.

@@

sp_bm <- sp_relation_plot2 %>%
  filter(x == "log_bm_std")
p_sp_bm <- sp_bm$gg 
names(p_sp_bm) <- sp_bm$y 
```
